---
title: "Project4_survival"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(warn=-1) # to prevent visualizing warnings in the report
ifelse("survival" %in% rownames(installed.packages()), NA, install.packages("survival"))
ifelse("survminer" %in% rownames(installed.packages()), NA, install.packages("survminer"))
ifelse("ranger" %in% rownames(installed.packages()), NA, install.packages("ranger"))
ifelse("ggplot2" %in% rownames(installed.packages()), NA, install.packages("ggplot2"))
ifelse("dplyr" %in% rownames(installed.packages()), NA, install.packages("dplyr"))
ifelse("ggfortify" %in% rownames(installed.packages()), NA, install.packages("ggfortify"))
ifelse("coin" %in% rownames(installed.packages()), NA, install.packages("coin"))
```

```{r libraries, include=FALSE}
library(survival)
library(survminer)
library(ranger)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(coin)
```

## 1) EDA
There are 6 variables in the dataset:
1) futime:	survival or censoring time
2) fustat:	censoring status (have people survived or not)
3) age:	in years
4) resid.ds:	residual disease present (1=no,2=yes)
5) rx:	treatment group
6) ecog.ps:	ECOG performance status (1 is better, see reference). The ECOG Scale of Performance Status describes a patient’s level of functioning in terms of their ability to care for themself, daily activity, and physical ability (walking, working, etc.) (from 0	- Fully active to 5 - Dead).

```{r descr}
ovarian
str(ovarian)
plot(ovarian)
```

All the variables are of numeric type. There are no NAs then.

```{r bp1}
bp1 <- ggplot(ovarian, aes(x=age)) +
  geom_histogram()
bp1

mean(ovarian$age)
median(ovarian$age)
```

Lets preprocess our dataset and create some factor variables.
As there are no obvious visible groups we will split all patients in 2 groups by age: 
less than 60 (mature people) and more than 60 years old (old people). These groups we will save in new "AG" variable. Also we factorize "resid.ds" and "rx" variable.

```{r ovar}
ovar <- mutate(ovarian, AG = ifelse((age < 60), "Less then 60", "More then60"),
               AG = factor(AG),
               resid.ds = factor(resid.ds,labels=c("residual desease present","residual desease not present")),
               rx = factor(rx,labels=c("first tr group","second tr group")))
```

The boxplot below shows that there is statistically significant difference in time before answer between two age groups. When patients are younger than 60 years, the time is bigger.

```{r}
bp2<-ggplot(ovar, aes(x=AG, y=futime, fill=rx)) +
  geom_boxplot()
bp2
```

As it is shown at the boxplot below there is statistically significant difference in time before answer between two age groups. When patients are younger than 60 years, the time is bigger.

```{r}
bp3<-ggplot(ovar, aes(x=AG, y=futime, fill=resid.ds)) +
  geom_boxplot()
bp3
```


## 2) Kaplan-Meier curves

Lets perform Kaplan-Meier curves analysys:

```{r}
km_fit <- survfit(Surv(futime, fustat) ~ 1, data=ovarian)
summary(km_fit, times = c(1,30,60,90*(1:10)))

autoplot(km_fit)
```

1)Lets have a look how rx (treatment group) variable change the curves:

```{r}
km_rx_fit <- survfit(Surv(futime, fustat) ~ rx, data=ovarian)
autoplot(km_rx_fit)
```

2)Lets have a look how resid.ds(residual disease present: 1=no,2=yes) variable change the curves:

```{r}
km_rx_fit <- survfit(Surv(futime, fustat) ~ rx, data=ovarian)
autoplot(km_rx_fit)
```

3)Lets have a look how age group change the curves. There is more obvious difference between the curves:

```{r}
km_AG_fit <- survfit(Surv(futime, fustat) ~ AG, data=ovar)
autoplot(km_AG_fit)
```


4)Lets have a look how the curves differ for 4 groups splitted by variables "AG" and "rx". We can see that the group "More then 60" with "Second tr" differ from other groups with more short survaval time.

```{r}
ovar_two_factors_rx <- ovar
ovar_two_factors_rx$hyb_fact <- paste(ovar_two_factors_rx$AG, ovar_two_factors_rx$rx)
km_AG_fit_rx <- survfit(Surv(futime, fustat) ~ hyb_fact, data=ovar_two_factors_rx)
autoplot(km_AG_fit_rx)
```

### 3) Groups that differ in survival

For age groups "AG" p-value is equal to 0.002903, then these groups are different.
```{r}
logrank_test(Surv(futime, fustat) ~ AG, data = ovar)
```

For resid.ds p-value is equal to 0.05444, then these groups are almost different.
```{r}
logrank_test(Surv(futime, fustat) ~ resid.ds, data = ovar)
```

For 4 groups splitted by AG + rx the p-value is also < 0.05.
```{r}
ovar_two_factors_rx$hyb_fact <- as.factor(ovar_two_factors_rx$hyb_fact)

logrank_test(Surv(futime, fustat) ~ hyb_fact, data = ovar_two_factors_rx)
```

### 4) Estimations of factors that affect the risk and risk ratio

Factors that affect the risk analysys:

```{r estimates}
cox <- coxph(Surv(futime, fustat) ~ ecog.ps + rx + age + resid.ds , data = ovar)
summary(cox)

cox_fit <- survfit(cox)
autoplot(cox_fit)

aa_fit <- aareg(Surv(futime, fustat) ~ ecog.ps + rx + age + resid.ds, data = ovar)
autoplot(aa_fit)
```

Risks ratio:

```{r 1}
fit.coxph <- coxph(Surv(futime, fustat) ~ ecog.ps + rx + age + resid.ds, data = ovar)
ggforest(fit.coxph, data = ovar)
```

